<style>
  .reveal pre {font-size: 12px;}
</style>

Introduction To Regular Expressions
========================================================
author: John S
date: 21 September 2015
font-family: 'Garamond'
transition: none

Agenda
===
- functions that work on strings
- regular expressions


Functions that work on strings
====

nchar() finds length(s) of string(s)


```{r}
x <- c("This","is","text")
nchar(x)
```

Functions that work on strings
====

paste() strings together

```{r}
paste("This","is","text",sep=" - ")
x <- c("This","is","text")
paste(x,collapse=" - ")
```

Functions that work on strings
====
substring() extracts elements between start and stop

```{r}
y <- substring("This",2,3)
paste(y,"!",sep="")
```

Functions that search on strings
====

grep(pattern,x) finds indices where text has first argument


```{r}
x <- c("This","is","text")
inds <- grep("i",x)
inds
x[inds]
```

Functions that search on strings
====

strsplit(x,split=pattern) splits x using delimeter

```{r}
x <- c("This - is - text.","So - is - this.")
y <- strsplit(x,split=" - ")
y
#turns the list into a vector
unlist(y)
```

Functions that search on strings
====

regexpr(pattern,x) finds first place pattern comes in x

```{r}
x <- c("This is text.","So is this.")
y <- regexpr("i",x)
y
c(y[1],y[2])
```

Functions that search on strings
====

gregexpr(pattern,x) finds all places a pattern comes in x

```{r}
x <- c("This is text.","So is this.")
y <- gregexpr("i",x)
y
```

Functions that search on strings
====

gsub(to find,to replace,x) replaces first argument with second.

```{r}
x <- c("20,123","21,113")
y <- gsub(",","",x)
y
as.numeric(y)
```

Regular expressions
====

Regular expressions are very flexible 'wildcards' that can be used in searches.

There are whole books about regular expressions, so this is an introduction! 

Can be used as "pattern" argument in previous functions.

Regular expressions
======
characters or numbers inside [ ] are treated as or

```{r}
# look for a or A in vector
grep("[aA]",c("ab","Ab","BB","BAa"))
```
[.] is any character and [0-9] is any number
```{r}
# look for the indices that contain integers in the vector
grep("[0-9]",c("ab","Ab","BB","BAa"))
```

Regular expressions
======
Some characters (.,$,\,etc) need to be preceeded by \\

```{r}
# look for any character in vector
grep(".",c("a.a","aa"))
# search fro special character is preceeded by \\
grep("\\.",c("a.a","aa"))
grep("\\$",c("$100","not paid"))
```

Regular expressions
=================
There are many special characters and there are quantifiers

\\s means space

\\s+[Aa] means at least 1 space followed by an A or a

```{r}
grep("\\s",c("a a","aa"))
#search for atleast one space followed by either A or a
grep("\\s+[Aa]",c("a A","aa"," a"))
```


Regular expressions
=================

Example: rich.html is a web page.

Want to use R to extract data from web page.

Want to make a data frame with names in one colums and net worth (in $Billions) in another.

This is lab on Fri. Next few pages help get you started. 


Regular expressions
====
```{r}
forbes <- readLines("~/rich.html")
length(forbes)
forbes[540:544]
```

Regular expressions
====

Text inside square brackets is like "or"

```{r}
#finds the gates or Gates in forbes
forbes <- readLines("~/rich.html")
inds <- grep("[gG]ates",forbes)
forbes[inds]
```

Regular expressions
====


```{r}
forbes <- readLines("~/rich.html")
inds <- grep("\\$",forbes)
head(forbes[inds])
tail(forbes[inds])
```


=======
```{r}
forbes <- readLines("~/rich.html")
#finding the dollar sign 
dollar.lines <- forbes[grep("\\$",forbes)]
dollar.lines <- dollar.lines[grep("\\sB",dollar.lines)]
tail(dollar.lines)
tail(forbes[grep("\\$[0-9,]+\\sB",forbes)])
```

=======
```{r}
# find all the names in the index right before the net income index
forbes <- readLines("~/rich.html")
names.lines <- forbes[grep("\\$[0-9,]+\\sB",forbes)-1]
head(names.lines)
tail(names.lines)
```

=====
```{r}
forbes <- readLines("~/rich.html")
names.lines <- forbes[grep("\\$[0-9,]+\\sB",forbes)-1]
#use a string split by finding <h3> in your vector and then split it 
#unlisting the stringsplit will give you a vector
head(unlist(strsplit(names.lines,split="<h3>"))[seq(from=2,to=200,by=2)])
#use gsub to replace everything that appears after the name with a blank 
```
