lecture14
========================================================
author: 
date: 

more about plyr
========================================================

In honor of world series, I'll start witha baseball example.
```{r,echo=F,cache=T}
library(plyr)
# dataset of careers of baseball players
# each row is player name identifier, year, team, and game information
data(baseball)
```
```{r,tidy=T,cache=T}
head(baseball)[,c(1,2,4,6:9)]
```

========
```{r,tidy=T,cache=T}
tail(baseball)[,c(1,2,4,6:9)]
dim(baseball)
```

2 functions: "transform" & "summarize"
=========

transform

```{r,tidy=T,cache=T}
# the trasnform function will take in a dataframe
# everything after the dataframe will be added as a column to the dataframe
standings <- data.frame(name=c("Mets","Royals"),W=c(90,95),L=c(72,67))
transform(standings,win.perc=round(W/(W+L),2),games=W+L)
```

2 functions: "transform" & "summarize"
=========

summarize

```{r,tidy=T,cache=T}
# summarise does similar things as transform
# But instead it creates a new data frame with just the new columns
standings <- data.frame(name=c("Mets","Royals"),W=c(90,95),L=c(72,67))
summarise(standings,win.perc=round(W/(W+L),2),games=W+L)
```

Goal: plot year as pro versus mean home run per at bat
====

1. Add column to baseball dataset with year as pro for each player, each year.
2. Calculate mean HR per at bat for possible values of year as pro
3. Make CI around that mean too

When doing something complicated, I usually write simpler functions first.

Add a column to baseball with year in league
========================================================

Get data for one player and try it

```{r,cache=T}
carl.y <- subset(baseball,id=="yastrca01")
head(carl.y)[,c(1,2,4,6:9)]
```


Add a column to baseball with year in league
========================================================

```{r,cache=T}
# adding th column for the pro years
carl.y <- transform(carl.y,pro.year=year-min(year)+1)
names(carl.y)
```

Use ddply to do it for each player (id)
=============

```{r,cache=T,tidy=T,fig.width=10}
par(cex=1.5)
# here we give ddply
  # dataframe- baseball
  # splitting criteria - id
  # function - transform
  # the extra arguments to function - transform taking in pro.year...
baseball <- ddply(baseball,.(id),transform,pro.year=year-min(year)+1)
# use vectors to plot
plot(1:length(table(baseball$pro.year)),as.vector(table(baseball$pro.year)),xlab="year",ylab="n",type="b")
```

Plot shows that people come in to the league and then don't play for a year but then come back for 3rd year. Hence ups and drops in the data. This data maxes for year 11 because the dataset we are dealing with would be a subset of the whole data. So in this case, we only have people data for people who max stayed in the league for 11 years. 

Get mean HR per AB (by pro year)
=============
First do it for two players

```{r,cache=T}
# take 2 players
two.carls <- subset(baseball,(id=="yastrca01")|(id=="fiskca01"))
```

How do we estimate the mean of a ratio? Let's use maximum likelihood.

-calculate the home run average for each player each year

-then apply **ddply** to get the average of the ratios

**OR**

-calculate the total home run for each pro.year

-calculate the total at bat for each pro.year

-Take the ratio of those totals for each pro.year

Calculate mean ratio
=========
```{r,cache=T,fig.width=10}
par(cex=2)
hr.per.ab.1 <- ddply(two.carls,.(pro.year),summarize,hr.rate=sum(hr)/sum(ab))
plot(hr.per.ab.1,col="red")
points(hr.per.ab.1,pch=16,col="blue")
```

Add n and se
====
```{r,cache=T}
# n= length of the id
# making a data frame with pro.year, home run rate and n
hr.rate.sum <- ddply(two.carls,.(pro.year),summarize,hr.rate=sum(hr)/sum(ab),n=length(id))
# adding the standard error
# getting the standard error of each p hats
hr.rate.sum <- transform(hr.rate.sum,se=sqrt(hr.rate*(1-hr.rate)/n))
head(hr.rate.sum)
```

Do it for all the data
====

```{r,cache=T,tidy=T}
# same for all data
hr.rate.sum <- ddply(baseball,.(pro.year),summarize,hr.rate=sum(hr)/sum(ab),n=length(id))
hr.rate.sum <- transform(hr.rate.sum,se=sqrt(hr.rate*(1-hr.rate)/n))
head(hr.rate.sum)
```


Make CIs
====

```{r,cache=T,tidy=T}
# now getting the CIs
hr.rate.sum <- transform(hr.rate.sum,lower=hr.rate-qt(.975,df=n-1)*se,upper=hr.rate+qt(.975,df=n-1)*se)
# replace NAN with 0
hr.rate.sum[hr.rate.sum=="NaN"] <- 0
head(hr.rate.sum)
```

Make a nice plot!
=======
```{r,cache=T,tidy=T,fig.width=12,eval=F}
par(cex=2)
ylim <- range(c(hr.rate.sum$lower,hr.rate.sum$upper))
plot(hr.rate.sum$pro.year,hr.rate.sum$hr.rate,ylim=ylim,
     xlab="pro year",ylab="HR rate (95% CI)",type="n",axes=F)
lines(hr.rate.sum$pro.year,hr.rate.sum$hr.rate,lwd=2)
lines(hr.rate.sum$pro.year,hr.rate.sum$upper,col="red",lwd=3)
lines(hr.rate.sum$pro.year,hr.rate.sum$lower,col="red",lwd=3)
axis(1)
axis(2)
```

Depending on what you are look for. You can cut the plot off at the early part of the carrer of the player

Make a nice plot! (not this one!)
=======
```{r,cache=T,tidy=T,fig.width=12,echo=F}
par(cex=2)
ylim <- range(c(0,hr.rate.sum$lower,hr.rate.sum$upper))
plot(hr.rate.sum$pro.year,hr.rate.sum$hr.rate,ylim=ylim,
     xlab="pro year",ylab="HR rate (95% CI)",type="n",axes=F)
lines(hr.rate.sum$pro.year,hr.rate.sum$hr.rate,lwd=3)
lines(hr.rate.sum$pro.year,hr.rate.sum$upper,col="red",lwd=3)
lines(hr.rate.sum$pro.year,hr.rate.sum$lower,col="red",lwd=3)
axis(1)
axis(2)
```

Make a nicer plot!
=======
```{r,cache=T,tidy=T,fig.width=12,echo=F}
par(cex=2)
hr.rate.sum <- subset(hr.rate.sum,(pro.year<21))
ylim <- range(c(0,hr.rate.sum$lower,hr.rate.sum$upper))
plot(hr.rate.sum$pro.year,hr.rate.sum$hr.rate,ylim=ylim,
     xlab="pro year",ylab="HR rate (95% CI)",type="n",axes=F)
lines(hr.rate.sum$pro.year,hr.rate.sum$hr.rate,lwd=3)
lines(hr.rate.sum$pro.year,hr.rate.sum$upper,col="red",lwd=3)
lines(hr.rate.sum$pro.year,hr.rate.sum$lower,col="red",lwd=3)
axis(1,at=c(1,5,10,15,20))
axis(2)
```


