<style>
  .reveal pre {font-size: 12px;}
</style>

Reading and Writing Data
========================================================
author:
date:
transition: none
font-family: 'Garamond'
  
In Previous Episodes
==========  
- Seen functions to load data in passing (e.g. readLines())
- Learned about string manipulation and regular expressions

Agenda
==========  
- Getting data into and out of the system when it's already in R format
- Import and export when the data is already very structured and machine-readable
- Dealing with less structured data
- Web scraping

Reading Data from R
==========  
- You can load and save R objects
+ R has its own format for this, which is shared across operating systems
+ It's an open, documented format if you really want to pry into it
- `save(thing, file="name")` saves `thing` in a file called `name` (conventional extension: `rda` or `Rda` or `Rdata`)
- `load("name")` loads the object or objects stored in the file called `name`, _with their old names_

Example
=====
```{r,echo=TRUE}
gmp <- read.table("gmp.dat")
gmp$pop <- round(gmp$gmp/gmp$pcgmp)
save(gmp,file="gmp.Rda")
rm(gmp)
exists("gmp")
load(file="gmp.Rda")
colnames(gmp)
```

==========  
- We can load or save more than one object at once; this is how RStudio will load your whole workspace when you're starting, and offer to save it when you're done
- Many packages come with saved data objects; there's the convenience function `data()` to load them
```{r}
# methods of applied statistics
data(cats,package="MASS")
summary(cats)
```

Non-R Data Tables
==========  
- Tables full of data, just not in the R file format
- Main function: `read.table()`
+ Presumes space-separated fields, one line per row
+ Main argument is the file name or URL
+ Returns a dataframe
+ Lots of options for things like field separator, column names, forcing or guessing column types, skipping lines at the start of the file...
- `read.csv()` is a short-cut to set the options for reading comma-separated value (CSV) files
+ Spreadsheets will usually read and write CSV

Example: Reading data can be tricky
==========  
```{r,cache=TRUE}
# because it has a character in it, it will turn it all into a character
data <- read.csv("huh.txt")
data
#can not do this because of character
data$speed/data$distance
```
huh

Example: Reading data can be tricky
==========  
```{r,cache=TRUE}
is.numeric(data$speed)
is.numeric(data$distance)
#distance is a factor variable
#order of factor
data$distance
#use levels(data$distance) shows that there exists a space
```

Example: Reading data can be tricky
==========  
```{r,cache=TRUE}
# taking a factor and turning it to a numerics
as.numeric(data$distance)
# make it a character first and then turn them into numbers
data$distance <- as.numeric(as.character(data$distance))
# warning: NA are introduced by the conversion
data$distance
# but now you can divide speed by distance
data$speed/data$distance
```

Writing Dataframes
==========  
- Counterpart functions `write.table()`, `write.csv()` write a dataframe into a file
- Drawback: takes a lot more disk space than what you get from `load` or `save`
- Advantage: can communicate with other programs

Less Friendly Data Formats
==========  
- The `foreign` package on CRAN has tools for reading data files from lots of non-R statistical software
- Spreadsheets are special

Spreadsheets Considered Harmful
==========  
- Spreadsheets look like they should be dataframes
- Real spreadsheets are full of ugly irregularities
+ Values or formulas?
+ Headers, footers, side-comments, notes
+ Columns change meaning half-way down
+ Whole separate programming languages apparently intended to mostly to spread malware

Spreadsheets, what to do
==========  
- Save the spreadsheet as a CSV; `read.csv()`
- Save the spreadsheet as a CSV; edit in a text editor; `read.csv()`
- Use `read.xls()` from the `gdata` package
+ Tries very hard to work like `read.csv()`, can take a URL or filename
+ Can skip down to the first line that matches some pattern, select different sheets, etc.
+ You may still need to do a lot of tidying up after


Semi-Structured Files, Odd Formats
==========  
- Sometimes non-tabular arrangement
- Generally, write function to read in one (or a few) lines and split it into some nicer format
+ Regular expressions can help
+ Best to write functions to do this
- Example: Actigraph accelerometer

Actigraph accelerometer example
==========
First few lines of dataset:
```
Created By ActiGraph GT3X+ ActiLife v6.7.0 Firmware v2.4.0 
date format M/d/yyyy at 80 Hz  Filter Normal ---
Start Time 09:15:00
Start Date 3/5/2012
Epoch Period (hh:mm:ss) 00:00:00
Download Time 10:33:16
Download Date 3/5/2012
Current Memory Address: 0
Current Battery Voltage: 4.22     Mode = 12
--------------------------------------------------
0.924,0.023,-0.32
0.927,0.026,-0.32
0.93,0.026,-0.32
```


Reader function
==========

```{r}
# function that reads in file
# input the path to file
read.actigraph.GT3X <- function(file.name.and.path)
{
  # read top 10 lines
  # read header and start time
  data <- readLines(paste(file.name.and.path),n=10)
  # read in the third/fourth element and split it and keep the part you want
  start.time <- data[3]
  start.time <- (strsplit(start.time,split=" ")[[1]][3])
  start.date <- data[4]
  start.date <- (strsplit(start.date,split=" ")[[1]][3])
  start.time <- as.POSIXlt(strptime(paste(start.date,start.time),
                                    "%m/%d/%Y %H:%M:%S"))

  # read acceleration data and add timestamps
	data <- read.csv(file.name.and.path,skip=10,header=F)
  n <- dim(data)[1] # number of rows in data
  	full.data <- 
  	  data.frame(time = start.time + (0:(n-1))/80,data)
  	
  return(full.data)
}
```

How'd I figure out how to do this?



Building data from the web
==========  
- Can do this with regular expressions like in lab
- Easier way is to use a package (e.g. XML)
- Example: Ironman data

Example: Ironman data
===========
- Want to build a dataset from the results tables lie at
http://www.ironman.com/triathlon/events/americas/ironman/mont-tremblant/results.aspx?p=1&race=monttremblant&rd=20150816&agegroup=Pro&sex=M&ps=20#axzz3BcHJd6dJ
- XML is a package that has tools to find tables on web pages

Example: Ironman data
==========
```{r}
library(XML)
junk <- "http://www.ironman.com/triathlon/events/americas/ironman/mont-tremblant/results.aspx?p=1&race=monttremblant&rd=20150816&agegroup=Pro&sex=M&ps=20#axzz3BcHJd6dJ"
# uses the htmlParse
junk.doc <- htmlParse(junk)
# finds all the tables in the data set
tab <- readHTMLTable(junk.doc)$eventResults
head(tab[,-c(2:5,10)])
```

I want all the data!
===
(See R console...)

Summary
==========  
- Loading and saving R objects is very easy
- Reading and writing dataframes is pretty easy
- Extracting data from unstructured sources is possible
